---
title: "IFRthemes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{IFRthemes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
# knitr::opts_chunk$set(
#   collapse = TRUE,
#   comment = "#>"
# )
```

```{r setup, message=FALSE}
library(IFRthemes)
library(ggplot2)
library(dplyr)
```

## Introduction

All IFR plots should be produced in ggplot2 using a standard theme and consistent colour palette. The `IFRthemes` package allows the user to apply a custom IFR theme and two different colour palettes depending on the number of colours required.

## Installation

You can install the package from GitHub using the `remotes::install_github` function. Using *build_vignettes = TRUE* allows you to access this vignette.

```{r load, eval=FALSE}
remotes::install_github("InStreamFisheries/IFRthemes",
                        build_vignettes = TRUE, 
                        force = TRUE)
library(IFRthemes)
```

Once the package is installed, you can access this vignette via the following command:

```{r, eval = FALSE}
browseVignettes("IFRthemes")
```

## Custom Theme

To use the IFR theme, simply add *+ theme_ifr()* to a ggplot object.

```{r, fig.height=5, fig.width=7}
ggplot(mtcars, aes(mpg, disp)) + 
  geom_point(size = 3) + 
  theme_ifr()
```

The theme also formats panels.

```{r, fig.height=4, fig.width=7}
ggplot(mtcars, aes(mpg, disp)) + 
  geom_point(size = 3) + 
  facet_wrap(~cyl) +
  theme_ifr()
```

### Changing Plot Text Size

There is one argument to the theme: base_size, which scales the size of plot text. The function defaults to 16, which sizes nicely for 6 inch figures. In some cases (e.g., very large plots with lots of faceting) you may need to change the text size. The example below shows how to apply this argument.

```{r, fig.height=4, fig.width=7}
ggplot(mtcars, aes(mpg, disp)) + 
  geom_point(size = 3) + 
  facet_wrap(~cyl) +
  theme_ifr(base_size = 10)
```

### Overwriting the Theme

It's easy to adjust the theme. Below the theme is overwritten to angle the x-axis text.

```{r, fig.height=5, fig.width=7}
ggplot(mtcars, aes(mpg, disp)) + 
  geom_point(size = 3) + 
  theme_ifr() +
  theme(legend.position = c(1, 1),
        axis.text.x = element_text(angle = 45, 
                                   hjust = 1))
```

## Colours

We have created two colour pallettes for use within ggplot2. The first colour palette (`scale_colour_ifr` or `scale_fill_ifr`) consists of six colours that are complementary to IFR's logo, which should be used for all plots with six or less colours. The second palette (`scale_colour_ifr10` or `scale_fill_ifr10`) is a larger palette of 10 colourblind-friendly colours that is used whenever there are more than six colours in a plot (typically rare).

The custom palettes are applied like any other ggplot colour palette.

```{r, fig.height=5, fig.width=7}
ggplot(mtcars, aes(as.factor(carb), fill = as.factor(carb))) + 
  geom_bar() + 
  scale_fill_ifr() +
  theme_ifr()
```

```{r, fig.height=5, fig.width=7}
ChickWeight %>%
   mutate(Chick = as.factor(Chick)) %>%
   group_by(Chick) %>%
   summarise(Mean = mean(weight),
            .groups = 'drop') %>%
   slice(1:10) %>%
   ggplot(., aes(x = Chick, y = Mean, fill = Chick)) +
   geom_col() +
   scale_fill_ifr10() +
   theme_ifr() 
```

### Viewing IFR Colours

A simple function (`show_colour_ifr`) was created to display the six custom colours and allow the user to select colours for a custom plot. For example, when overlaying multiple plot types or adding lines to plots, the user may want to select colours without applying the full palette.

```{r, fig.height=4, fig.width=5}
show_ifr_colours()
```

```{r, fig.height=5, fig.width=7}
ggplot(mtcars, aes(mpg, disp)) + 
  geom_point(size = 3) + 
  theme_ifr() +
  geom_vline(xintercept = 20, 
             size = 3,
             color = "#a40c01") +
  geom_hline(yintercept = 300,
             size = 3,
             color = "#104692")
```

## Plotting Best Practices:

1. Use ggplot2 whenever possible
2. Save plots to file using ggsave:
   * Resolution should be set to 600 dpi
   * Scale the plot size using the inches argument
   * Typically save plots as 6 inches wide to fit the width of a 8.5 x 11 sheet
   * ggsave("plot.png", plot, height = 6, width = 6, units = "in", dpi = 600)
3. Do not use titles on plots
4. Always use legends rather than explaining the colours in the caption 
5. Gridlines can be added if necessary for interpretation/clarity
   * E.g., panel.grid.major.x = element_line(size = 0.25)
6. Use commas or "K" if you numbers in the thousands as y-axis labels
   * E.g., using the scales package: scale_y_continuous(label = comma)
7. Angle your x-axis when necessary
   * theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
8. Don't use an x-axis label for time series data (e.g., dates or years)
9. Use the abbreviated month when plotting dates:
   * E.g., "2020-Jan-15", or "Jan-15" (if no year is required)

